---
title: "Getting and Cleaning Data"
author: "Luis Moreno"
date: "9/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Note: the `echo = FALSE` parameter is added to the code chunk to prevent printing of the R code.

## Reading Excel files
```{r}
fileURL <- "https://data.baltimorecity.gov/api/views/dz54-2aru/rows.xlsx?accessType=DOWNLOAD"
download.file(fileURL, destfile = "./data/cameras.xlsx", method = "curl")
dateDownloaded <- date()
```

### read.xlsx()
This function allows the user to read data from an Excel worksheet. It provides the conveniency of read.table by borrowing from its signature.  

### Main arguments for read.xlsx function
```{reading_xlsx, include = TRUE}
library(xlsx)
str(read.xlsx)
```
  
**file**: the path to the file to read.  
**sheetIndex**: a number representing the sheet index in the workbook.  
**sheetName**: a character string with the sheet name.  
**rowIndex**: a numeric vector indicating the rows you want to extract.  
**colIndex**: a numeric vector indicating the cols you want to extract.  
**head**: a logical value indicating whether the first row corresponding to the first element of the *rowIndex* vector contains the names of the variables

```{r}
colIndex <- 2:3
rowIndex <- 1:4
cameraDataSubset <- read.xlsx("./data/cameras.xlsx", sheetIndex = 1,
                              colIndex = colIndex, rowIndex = rowIndex)
cameraDataSubset
```

### Further Notes  
**write.xlsx**: it will write out an Excel file with similar arguments.  
**read.xlsx2**: it is faster than read.xsls but for reading subsets of rows is more unstable.  
**XLConnect**: this package has more options for writing and manipulating Excel files.  
**XLConnect vignette**: is a good place to start for that package.  
**Advise**: it is better to store data into a database or in comma separated files (.csv) or tab separated files (.tab/.txt).  

## Reading XML files
### XML
XML is the abbreviation for "Extensible Markup Language". It is frequently used to store *structured data*, particularly in internet applications.  
Its components are:  
