---
title: "Week 1. GACD Quiz"
author: "Luis Moreno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1  

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:   https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv  
How many properties are worth $1,000,000 or more?  
```{r, eval = FALSE}
library(readr)
us_comminities <- "./data/us_communities.csv"
df <- data.frame(read_csv(us_comminities))
VAL <- df$VAL
VAL <- na.omit(df$VAL)length(mio_val)
mio_val <- subset(VAL, VAL == "24")
length(mio_val)
```

### Question 2  
Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?   
**Answer**: Each variable in the tidy data set has been transformed to be interpretable.  

### Question 3  
Download the Excel spreadsheet on Natural Gas Aquisition   Program here: https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx  
Read rows 18-23 and columns 7-15 into R and assign the result to a variable called "dat".    
What is the value of: sum(dat$Zip*dat$Ext,na.rm=T)  
```{r, eval = FALSE}
library(readxl)
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
download.file(fileURL, destfile = "./data/nat_gas_adq.xlsx")
xlsx_file_path <- "./data/nat_gas_adq.xlsx"
dat <- read_excel(xlsx_file_path, range = "G18:O23")
sum(dat$Zip*dat$Ext, na.rm = TRUE)
```

### Question 4   
Read the XML data on Baltimore restaurants from here:   https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml  
How many restaurants have zipcode 21231?  
```{r, eval = FALSE}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
download.file(fileURL, destfile = "./data/baltimore_restaurants.xml")
library(XML)
file_path <- "./data/baltimore_restaurants.xml"
doc <- xmlTreeParse(file_path, useInternalNodes = TRUE)
rootNode <- xmlRoot(doc)
zip_codes <- xpathSApply(rootNode, "//zipcode", xmlValue)
length(subset(zip_codes, zip_codes == "21231"))
```

### Question 5  
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:  
https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv  
using the fread() command load the data into an R object "DT" The following are ways to calculate the average value of the variable "pwgtp15" broken down by sex.  
Using the data.table package, which will deliver the fastest user time?    

```{r, eval = TRUE}
fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
library(readr)
library(data.table)
file_path <- "./data/american_comm_survey.csv"
DT = fread(file_path)
system.time(mean(DT$pwgtp15, by = DT$SEX))
system.time(tapply(DT$pwgtp15, DT$SEX, mean))
system.time(DT[, mean(pwgtp15), by = SEX])
system.time(sapply(split(DT$pwgtp15, DT$SEX), mean))
system.time(mean(DT[DT$SEX == 1, ]$pwgtp15)) + system.time(mean(DT[DT$SEX == 2, ]$pwgtp15))

system.time(tapply(DT$pwgtp15,DT$SEX,mean))
system.time(mean(DT[DT$SEX==1,]$pwgtp15), mean(DT[DT$SEX==2,]$pwgtp15))
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
system.time(DT[,mean(pwgtp15),by=SEX])
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
```
**ANSWER: DT[,mean(pwgtp15), by=SEX]** 