---
title: "GACD. Week 4 Quiz"
author: "Luis Moreno"
date: "1/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Loading Packages}
library(tidyverse)
```

## Question 1  
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf 

Apply `strsplit()` to split all the names of the data frame on the characters "wgtp". What is the value of the 123 element of the resulting list?

```{r Answer 1}
# Download data file and store it
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv", destfile = "./data/american_comm_survey.csv")

# Set the working directory
setwd("/Users/luism/Documents/R/getting-cleaning-data/data")

# Load the downloaded data into R Studio
dat <- read_csv("./data/american_comm_survey.csv")

# Create a vector containing the column names of the data
dat_col <- colnames(dat)

# Split the names of the dataframe on the characters "wgtp"
splitted <- strsplit(dat_col, "wgtp")
splitted[123]
```

## Question 2  
Load the Gross Domestic Product data for the 190 ranked countries in this data set: 

 https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv 

Remove the commas from the GDP numbers in millions of dollars and average them. What is the average? 

Original data sources: 

http://data.worldbank.org/data-catalog/GDP-ranking-table
```{r Answer 2}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv",
              destfile = "./data/countries_gdp.csv")

dat <- read_csv("./data/countries_gdp.csv",
                col_names = c("country_code",
                              "gdp_rank",
                              "x3",
                              "economy",
                              "gdp_usdM"),
                skip = 5)

# Clean the CSV file by subsetting and creating a new data-frame
c12 <- dat[1:190, 1:2]
c34 <- dat[1:190, 4:5]
dat2 <- tibble(c12, c34)

# Substitute commas with blank space of a column
gdp_wo_commas <- gsub(",", "", dat2$gdp_usdM)

# Calculate the average of a non-numeric value
mean(as.numeric(gsub(",", "", gdp_wo_commas)))
```

## Question 3  
In the data set from Question 2 what is a regular expression that would allow you to count the number of countries whose name begins with "United"?  

Assume that the variable with the country names in it is named countryNames.  

How many countries begin with United? 
```{r Answer 3}
# Create a vector with the country names
countryNames <- dat2$economy

# Number of coincidences starting with a character 
## Countries starting with "United" Method 1
length(grep("*United",countryNames))
## Countries starting with "United" Method 2
length(grep("^United",countryNames))
```

## Question 4  
Load the Gross Domestic Product data for the 190 ranked countries in this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv 

Load the educational data from this data set:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv

Match the data based on the country shortcode. Of the countries for which the end of the fiscal year is available, how many end in June?

Original data sources: 

http://data.worldbank.org/data-catalog/GDP-ranking-table  

http://data.worldbank.org/data-catalog/ed-stats  
```{r Answer 4}
# Load the first CSV file
gdp_ranks <- read_csv("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv", 
                      skip = 5,
                      col_names = c("country_code",
                                    "gd_rank",
                                    "x3",
                                    "economy",
                                    "gdp_usdM"))
# Clean the first CSV file
gr1 <- gdp_ranks[1:190, 1:2]
gr2 <- gdp_ranks[1:190, 4:5]
gdp_ranks <- tibble(gr1, gr2)

# Load the second CSV file
education <- read_csv("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv")

# Load the data.table package
library(data.table)

# Create the data.table objects
dt1 <- data.table(gdp_ranks)
dt2 <- data.table(education)

# Set the keys to sort them in ascending order
setkey(dt1, country_code)
setkey(dt2, CountryCode)

# Merge the datasets
merged_data <- merge(x = dt1, 
                     y = dt2,
                     by.x = "country_code",
                     by.y = "CountryCode")
# Create a variable where coincidences will be search
fiscal_info <- na.omit(object = merged_data$`Special Notes`)

# Count the number of coincidences
length(grep(pattern = "[J]une",
            x = fiscal_info))
```

## Question 5  
You can use the [quantmod](http://www.quantmod.com/) package to get historical stock prices for publicly traded companies on the NASDAQ and NYSE.  

Use the following code to download data on Amazon's stock price and get the times the data was sampled.

```{r}
library(quantmod)

amzn <- getSymbols("AMZN",
                  auto.assign=FALSE)

sampleTimes <- index(amzn)
class(sampleTimes)
```

How many values were collected in 2012? How many values were collected on Mondays in 2012?

```{r Answer 5}
# Load required packages
library(lubridate)

# Count the number of dates that start with "2012"
sum(
     grepl(pattern = "^2012",
           x = sampleTimes)
)

# Create a logical vector for 2012 dates
logical_date <- grepl(pattern = "^2012",
                      x = sampleTimes)

# Subset dates for the year 2012
twoK_twelve <- subset(sampleTimes, 
                              logical_date)

# Create a vector containing the day names
days_names <- wday(twoK_twelve,
                   label = TRUE)

# Count the number of Mondays
sum(
     grepl(pattern = "^Mon",
           x = days_names)
)
```

# Solution Key  

1.        "" "15"  
2.        377652.4  
3.        grep("^United",countryNames), 3  
4.        8  
5.        250, 47  