---
title: "GACD Week 2 Quiz"
author: "Luis Moreno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1. 
Instructions: Register an application with the Github API here https://github.com/settings/applications. Access the API to get information on your instructors repositories (hint: this is the url you want "https://api.github.com/users/jtleek/repos").  
Use this data to find the time that the datasharing repo was created.  
What time was it created?
```{r, eval = FALSE}
library(jsonlite)
URL <- "https://api.github.com/users/jtleek/repos"
jsonData <- fromJSON(URL)
names(jsonData)
df <- data.frame(jsonData)
library(dplyr)
ds <- filter(df, name == "datasharing")
names(ds)
ds <- select(ds, c("id", "name", "created_at", "updated_at"))
ds$created_at
```

### Question 2 and 3  
The **sqldf** package allows for execution of SQL commands on R data frames. We will use the sqldf package to practice the queries we might send with the dbSendQuery command in RMySQL.  
Download the American Community Survey data and load it into an R object called "acs"  
Which of the following commands will select only the data for the probability weights pwgtp1 with ages less than 50?  
```{r, eval = FALSE}
# Download the CSV file from the Web
file_URL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
download.file(file_URL, destfile = "./data/acs.csv", method = "curl")
# Read the downloaded file as a CSV and convert it to a data frame
file_path <- "./data/acs.csv"
library(readr)
acs <- read_csv(file_path)
acs <- data.frame(acs)
# Install the "sqldf" package, including H2
install.packages("sqldf", dep = TRUE)
# Load the "sqldf" package
library(sqldf)
# Select only the data for the probability weights "pwgtp1" with ages less than 50
sqldf("select pwgtp1 from acs where AGEP < 50")
# What is the equivalent function to "unique(acs$AGEP) in SQL?"
u <- data.frame(unique(acs$AGEP))
class(u)
head(u, 10)
u2 <- sqldf("select distinct AGEP from acs")
class(u2)
head(u2, 10)
head(u == u2, 5)
```
### Question 4  
How many characters are in the 10th, 20th, 30th and 100th lines of HTML from this page: http://biostat.jhsph.edu/~jleek/contact.html  
```{r, eval = FALSE}
con = url("http://biostat.jhsph.edu/~jleek/contact.html")
htmlCode = readLines(con)
close(con)
# Extract characters in the 10th, 20th, 30th and 100th lines
htmlCode[c(10, 20, 30, 100)]
# Count the number of characters of each line
nchar(htmlCode[c(10, 20, 30, 100)])
```
### Question 5  
Read this data set into R and report the sum of the numbers in the fourth of the nine columns.  
https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for  
Original source of the data: http://www.cpc.ncep.noaa.gov/data/indices/wksst8110.for  

(Hint this is a fixed width file format)  
```{r, eval = FALSE}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for", destfile = "./data/weekly_sst.txt", method = "curl")
library(readr)
file_path <- "./data/weekly_sst.txt"
columns <- c("week", "n1_sst")
df <- read_fwf(file_path, fwf_empty(file_path, col_names = c("week", "n1+2_SST", "n1+2_SSTA", "n3_SST", "n3_SSTA", "n34_SST", "n34_SSTA", "n4_SST", "n4_SSTA")))
```