---
title: "GACD. Week 3 Quiz"
author: "Luis Moreno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1  
The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv  

and load the data into R. The code book, describing the variable names is here:  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf  

Create a logical vector that identifies the households on greater than 10 acres who sold more than $10,000 worth of agriculture products. Assign that logical vector to the variable agricultureLogical. Apply the which() function like this to identify the rows of the data frame where the logical vector is TRUE.  

*which(agricultureLogical)*  

What are the first 3 values that result?  
```{r}
library(readr);library(dplyr)
file <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(file, destfile = "./data/us_comm.csv", method = "curl")
csv <- read_csv("./data/us_comm.csv")
agricultureLogical <- csv$ACR == 3 & csv$AGS == 6
head(which(agricultureLogical), 3)
```

### Question 2  
Using the jpeg package read in the following picture of your instructor into R  

https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg  

Use the parameter native=TRUE. What are the 30th and 80th quantiles of the resulting data? (some Linux systems may produce an answer 638 different for the 30th quantile)  
```{r}
# Install the jpeg Package
install.packages("jpeg")
library(jpeg)

# Set the URL of the jpeg file
jpeg_file <- "https://d396qusza40orc.cloudfront.net/getdata%2Fjeff.jpg"

# Download the jpeg file
download.file(jpeg_file, destfile = "./data/jeff_leek_photo.jpge", method = "curl")

# Read the JPEG file
jeff_leek <- readJPEG("./data/jeff_leek_photo.jpge", native = TRUE)

# Calculate the 30th and 80th quantile
quantile(jeff_leek, probs = c(0.3, 0.8))
```
### Question 3  
Load the Gross Domestic Product data for the 190 ranked countries in this data set:  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv  

Load the educational data from this data set:  

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv  

Match the data based on the country shortcode. How many of the IDs match? Sort the data frame in descending order by GDP rank (so United States is last). What is the 13th country in the resulting data frame?  

Original data sources:  

http://data.worldbank.org/data-catalog/GDP-ranking-table  

http://data.worldbank.org/data-catalog/ed-stats  
```{r, eval = FALSE}
# URL's of the files to be downloaded
url1 <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
url2 <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"

# Download the CSV files
download.file(url1, destfile = "./data/countries_gdp.xlsx", method = "curl")
download.file(url2, destfile = "./data/educational_data.csv", method = "curl")

# Read the downloaded CSV files
gdp_csv <- read_csv("./data/countries_gdp.csv")
edu_csv <- read_csv("./data/educational_data.csv")

# Return the column names for each data set
names(edu_csv)
names(gdp_csv)

# Convert the CSV files into data frames
gdp_df <- data.frame(gdp_csv)
edu_df <- data.frame(edu_csv)

# Count the number of matches
length(na.omit(merged_data$ranking))

# Sort the data frame in descending order by GDP rank
merged_ordered <- arrange(merged_data, desc(merged_data$ranking))

# Select the 13t country from the ordered data frame
merged_ordered[13, "Long.Name"]
```

### Question 4  
What is the average GDP ranking for the "High income: OECD" and "High income: nonOECD" group?  
```{r, eval = FALSE}
# Subset OECD High Income countries data
high_income_OECD <- subset(merged_ordered, merged_ordered$Income.Group == "High income: OECD")
# Calculate the mean of the ranking
mean(na.omit(high_income_OECD$ranking))

# Subset OECD Non High Income countries data
non_high_income_OECD <- subset(merged_ordered, merged_ordered$Income.Group == "High income: nonOECD")

# Calculate the mean of the ranking (omit NA values)
mean(na.omit(non_high_income_OECD$ranking))
```

### Question 5  
Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries are Lower middle income but among the 38 nations with highest GDP?  
```{r, eval = FALSE}
# Cut the GDP ranking into 5 separate quantile groups
ranking_quantiles <- mutate(merged_ordered, quantiles = ntile(merged_ordered$ranking, 5))
# Make a contingency table to show how many lower middle income countries are among the 38 nations with highest GDP
table(ranking_quantiles$Income.Group, ranking_quantiles$quantiles, useNA = "ifany")
```